---
title: "Example of Linear Regression in Effort Estimation"
author: "KUC"
date: Oct 2021
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Introduction

Example of a _linear regression_ model in Software Engineering, we will use *ISBSG* dataset.



## Read the ISBSG dataset

```{r Read file, message=FALSE}
library(foreign)
library(tidyverse)

isbsg <- read.arff("datasets/isbsg10teaser.arff")
df <- isbsg[, c("FS", "N_effort")]

# Change the names of headings
df <- df %>% rename( size = FS,
               eff = N_effort)

head(df)
```



Descriptive statistics of the dataset (actually a dataframe in R)

```{r df}
summary(df)
```


```{r}
cor(df$size, df$eff)  # calculate correlation between size and effort 
```

Based on the correlation, it seems that it will be difficult to properly estimate project using this model.

The simplest plot, size vs. effort.

```{r pressure, echo=FALSE}
plot(df)
```


Scatter plots can help visualize any linear relationships between the dependent (_effort_) variable and independent (predictor) variables, in this case the size.

```{r}
scatter.smooth(x=df$size, y=df$eff, main="Size ~ Effort")  # scatterplot
```


Boxplots

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
# Boxplot for 'size'
boxplot(df$size, main="Size")  #  ,sub=paste("Outliers: ", boxplot.stats(df$size)$out)
boxplot(df$eff, main="Effort")  # , sub=paste("Outliers: ", boxplot.stats(df$eff)$out) box plot for 'eff'
```


```{r}
par(mfrow = c(1, 2))  #n figures per row
size <- df$size
eff <- df$eff

hist(size, col = "blue", xlab = "size", ylab = "Probability", main = "Histogram of project Size")
lines(density(size, na.rm = T, from = 0, to = max(size)))
plot(density(size))
```


```{r}
library(e1071)
par(mfrow=c(1, 2))  # divide graph area in 2 columns

plot(density(df$size), main="Density Plot: Size", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(df$size), 2)))  # density plot for 'speed'
polygon(density(df$size), col="red")

plot(density(df$eff), main="Density Plot: Effort", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(df$eff), 2)))  # density plot for 'dist'
polygon(density(df$eff), col="red")
```



```{r}
linearMod <- lm(eff ~ size, data=df)  # build linear regression model on full data
print(linearMod)
```

```{r}
modelSummary <- summary(linearMod) 
print(modelSummary)
```


## Proposed Exercises

1. As exercise, the validation of the model should be done using a testing dataset, i.e., unseen data, to so we need to divide the dataset into training and testing, create the model using training data and validate it the remaining data.

2. As the data is quite skew and it does not follow a normal distribution, a possible solution is to transform the data using logarithms and repeat the process of learning and validation. 
